The reference model approach in feature selection problems \vspace{.1in}

### 2019-07-15 paper/ref_approach.tex
unifed notation equations sec 2 and 2.1

A good predictive model is able to filter part of the noise present in the data. The noise is the main source of the instability and tends to obscure the relevance of the covariates in relation to the target variable of interest. Indeed, let us consider the following explanatory example taken from \cite{paper:projpred}. The data generation mechanism is
\begin{alignat}{2} \label{eq:simulated_data}
     &f\sim N(0,1) && \nonumber \\ 
     Y|&f\sim N(f,1) && \\
     X_{j}|&f \overset{iid}{\sim} N(\sqrt{\rho}f,1-\rho) \quad &&j=1,..,k \nonumber \\
     X_{j}|&f \overset{iid}{\sim} N(0,1) &&j=k+1,..,p \nonumber,
\end{alignat}
where $f$ is the latent variable of interest of which $Y$ is a noisy observation. The first $k$ covariates are strongly related to the target variable $Y$ and correlated among themselves. Precisely, $\rho$ is the correlation among any pair of the first $k$ covariates, whereas $\sqrt{\rho}$ and $\sqrt{\rho}/2$ are the level of correlation between any relevant covariate and respectively $f$ and $Y$. If we had an infinite amount of observations, the sample correlation would be equal to the true correlation, though, even in this ideal asymptotic regime, the correlation computed would be still a biased indicator of the true relevance of each variable with respect to the latent variable of interest due to the intrinsic noisy nature of $Y$.

When using a reference model, we first model $f$ using all the covariates $\{X_{j}\}_{j=1}^{p}$ taking into account that we observe noisy $Y$. If our model is good, we are able to describe $f$ better than the noisy $Y$, which imporves the accuracy in the estimation of the relevance of the covariates.  Figure \ref{fig:correlation} shows the scatter plot of the absolute value of the sample correlation of each feature with the point prediction by a reference model, in this case see model \eqref{eq:ref_mod}, against the absolute value of the sample correlation of each feature with the noisy observations $\{y\}_{i=1}^{n}$. Looking at the marginal distributions, we can see that using a reference model to filter out noise in the data the two groups of features (relevant and non-relevant) can be better distinguished than when the correlation is computed using directly the observed noisy data.

@@

% In Figure \ref{fig:correlation} the reference model filters the noise.
% This simple idea is what happens every time a reference model is used, as it is evident for example in the preconditioning by \cite{paul2008preconditioning} or in the draw-by-draw projection for GLMs by \cite{paper:original_proj}.\\
% Above the correlation was used to estimate the relevance of a covariate.  
As an another example, let us consider a candidate submodel in a feature selection process and let us refer with $\pi$ to the parameter distribution of the submodel and with $q_{\pi}(\tilde{y})$ to the induced predictive distribution. We would like to choose $\pi$ in order that the latter maximises some predictive performance utility, as the expected log-predictive density (elpd) defined as:
\
\begin{equation}\label{eq:elpd}
\text{elpd}[q_{\pi}]=\int \text{log}\,q_{\pi}(\tilde{y})p_{t}(\tilde{y})d\tilde{y} 
\end{equation}
If we refer to the posterior predictive distribution of a reference model with $p_{\text{ref}}(\tilde{y}|D)$, where $D$ stands for the data on which we condition on, we can approximate \eqref{eq:elpd} using $p_{\text{ref}}(\tilde{y}|D)$ instead of the true, unknown, data generation mechanism $p_{t}(\tilde{y})$. The maximisation of the elpd using the reference model predictive distribution is equivalent to the minimisation of the KL divergence between the reference model predictive distribution and the submodel one:
\
\begin{equation} \label{proj_as_filter}
\underset{\pi}{\text{max}} \; \int \text{log}\,q_{\pi}(\tilde{y})p_{\text{ref}}(\tilde{y}|D)d\tilde{y} \leftrightarrow \underset{\pi}{\text{min}} \; \text{KL}[p_{\text{ref}}(\tilde{y}|D)||q_{\pi}(\tilde{y})] 
\end{equation}
The term on the right-hand side of Equation \eqref{proj_as_filter} describes what is referred as the projection of the predictive distribution, which is the general idea behind the projection predictive approach. Again the reference model is acting as a filter, or substitute, on data.

@@

As a motivational example to the study of the consequences of the use of a reference model in feature selection problems, we show its application in the form of the projection predictive approach, which uses the reference model to project the posterior samples in the submodels. The \texttt{R}-package \texttt{projpred} (available at \url{https://CRAN.R-project.org/package=projpred}) implements the projection in case of submodels in the family of the generalised linear models and, additionally, it provides a framework to choose the optimal submodel size through a predictive utility comparison between the submodels and the reference model. 
\\
We now summarise the workflow of the projection predictive approach in the particular case of the draw-by-draw projection \cite[original formulation by][]{paper:original_proj}, following \cite{paper:projpred}. Assume to have $n$ statistical units with target values $\{y_{i}\}_{i=1}^{n}$ and a set of observed covariates, the main steps are the following:
\begin{enumerate}
\item Devise and fit a reference model. Let $\{\boldsymbol{\theta}_{*}^{s}\}_{s=1}^{S}$ be the set of S posterior samples of the vector parameter of the model.
\item Rank the covariates using a heuristics and consider as candidate submodels the nested ones, starting to include the best ranked feature. The submodel are naturally identified by their model size.
\item For each submodel, project the reference model samples as follows:
\begin{equation}\label{eq:draw_by_draw}
\boldsymbol{\theta}_{\perp}^{s} = \underset{\boldsymbol{\theta^{s}}\in\Theta}{\text{argmin}} \frac{1}{n}\sum_{i=1}^{n}\text{KL}\;[p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})\,||\,p(\tilde{y}_{i}|\boldsymbol{\theta}^{s})] \quad s=1,..,S
\end{equation}
where $p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})$ stands for the model distribution of the reference model with parameters fixed at the sample value $\boldsymbol{\theta}_{*}^{s}$ and conditioning on all the covariates related to the statistical unit (identified by the subscript $i$ in the variable $\tilde{y}_{i}$), whereas $p(\tilde{y}_{i}|\boldsymbol{\theta}^{s})$ the same but for the submodel. The projected samples correspond to the submodel parameter distribution.
\item For each submodel (size), test the predictive performance.
\item Choose the smallest submodel (size) that is sufficiently close to the reference model prediction utility score.
\end{enumerate}

>>>>

A good predictive model is able to filter part of the noise present in the data. The noise is the main source of the instability and tends to obscure the relevance of the covariates in relation to the target variable of interest. Indeed, let us consider the following explanatory example taken from \cite{paper:projpred}. The data generation mechanism is
\begin{alignat}{2} \label{eq:simulated_data}
     &f\sim N(0,1) && \nonumber \\ 
     Y|&f\sim N(f,1) && \\
     X_{j}|&f \overset{iid}{\sim} N(\sqrt{\rho}f,1-\rho) \quad &&j=1,..,k \nonumber \\
     X_{j}|&f \overset{iid}{\sim} N(0,1) &&j=k+1,..,p \nonumber,
\end{alignat}
where $f$ is the latent variable of interest of which $Y$ is a noisy observation. The first $k$ covariates are strongly related to the target variable $Y$ and correlated among themselves. Precisely, $\rho$ is the correlation among any pair of the first $k$ covariates, whereas $\sqrt{\rho}$ and $\sqrt{\rho/2}$ are the level of correlation between any relevant covariate and respectively $f$ and $Y$. If we had an infinite amount of observations, the sample correlation would be equal to the true correlation, though, even in this ideal asymptotic regime, the correlation computed would be still a biased indicator of the true relevance of each variable with respect to the latent variable of interest due to the intrinsic noisy nature of $Y$.

When using a reference model, we first model $f$ using all the covariates $\{X_{j}\}_{j=1}^{p}$ taking into account that we observe noisy $Y$. If our model is good, we are able to describe $f$ better than the noisy $Y$, which improves the accuracy in the estimation of the relevance of the covariates.  Figure \ref{fig:correlation} shows the scatter plot of the absolute value of the sample correlation of each feature with the point predictions $\{\hat{y}\}_{i=1}^{n}$ of a reference model, in this case posterior predictive means of model \eqref{eq:ref_mod}, against the absolute value of the sample correlation of each feature with the noisy observations $\{y\}_{i=1}^{n}$. Looking at the marginal distributions, we can see that using a reference model to filter out noise in the data the two groups of features (relevant and non-relevant) can be better distinguished than when the correlation is computed using directly the observed noisy data.

@@

% In Figure \ref{fig:correlation} the reference model filters the noise.
% This simple idea is what happens every time a reference model is used, as it is evident for example in the preconditioning by \cite{paul2008preconditioning} or in the draw-by-draw projection for GLMs by \cite{paper:original_proj}.\\
% Above the correlation was used to estimate the relevance of a covariate.  
As an another example, let us consider a candidate submodel in a feature selection process and let us refer with $\pi$ to the parameter distribution of the submodel and with $q_{\pi}(\tilde{y})$ to the induced predictive distribution. We would like to choose $\pi$ in order that the latter maximises some predictive performance utility, as the expected log-predictive density (elpd) defined as:
\
\begin{equation}\label{eq:elpd}
\text{elpd}[q_{\pi}]=\int \text{log}\,q_{\pi}(\tilde{y})p_{t}(\tilde{y})d\tilde{y} 
\end{equation}
If we refer to the posterior predictive distribution of a reference model with $p(\tilde{y}|D)$, where $D$ stands for the data on which we condition on, we can approximate \eqref{eq:elpd} using $p(\tilde{y}|D)$ instead of the true, unknown, data generation mechanism $p_{t}(\tilde{y})$. The maximisation of the elpd using the reference model predictive distribution is equivalent to the minimisation of the KL divergence between the reference model predictive distribution and the submodel one:
\
\begin{equation} \label{proj_as_filter}
\underset{\pi}{\text{max}} \; \int \text{log}\,q_{\pi}(\tilde{y})p(\tilde{y}|D)d\tilde{y} \leftrightarrow \underset{\pi}{\text{min}} \; \text{KL}[p(\tilde{y}|D)||q_{\pi}(\tilde{y})] 
\end{equation}
The term on the right-hand side of Equation \eqref{proj_as_filter} describes what is referred as the projection of the predictive distribution, which is the general idea behind the projection predictive approach. Again the reference model is acting as a filter, or substitute, on data.

@@

As a motivational example to the study of the consequences of the use of a reference model in feature selection problems, we show its application in the form of the projection predictive approach, which uses the reference model to project the posterior samples in the submodels. The \texttt{R}-package \texttt{projpred} (available at \url{https://CRAN.R-project.org/package=projpred}) implements the projection in case of submodels in the family of the generalised linear models and, additionally, it provides a framework to choose the optimal submodel size through a predictive utility comparison between the submodels and the reference model. 
\\
We now summarise the workflow of the projection predictive approach in the particular case of the draw-by-draw projection \cite[original formulation by][]{paper:original_proj}, following \cite{paper:projpred}. Assume to have $n$ statistical units with target values $\{y_{i}\}_{i=1}^{n}$ and a set of observed covariates, the main steps are the following:
\begin{enumerate}
\item Devise and fit a reference model. Let $\{\boldsymbol{\theta}_{*}^{s}\}_{s=1}^{S}$ be the set of S posterior samples of the vector parameter of the model.
\item Rank the covariates using a heuristics and consider as candidate submodels the nested ones, starting to include the best ranked feature. The submodel are naturally identified by their model size.
\item For each submodel, project the reference model samples as follows:
\begin{equation}\label{eq:draw_by_draw}
\boldsymbol{\theta}_{\perp}^{s} = \underset{\boldsymbol{\theta^{s}}\in\Theta}{\text{argmin}} \frac{1}{n}\sum_{i=1}^{n}\text{KL}\;[p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})\,||\,q(\tilde{y}_{i}|\boldsymbol{\theta}^{s})] \quad s=1,..,S
\end{equation}
where $p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})$ stands for the model distribution of the reference model with parameters fixed at the sample value $\boldsymbol{\theta}_{*}^{s}$ and conditioning on all the covariates related to the statistical unit (identified by the subscript $i$ in the variable $\tilde{y}_{i}$), whereas $q(\tilde{y}_{i}|\boldsymbol{\theta}^{s})$ the same but for the submodel. The projected samples correspond to the submodel parameter distribution.
\item For each submodel (size), test the predictive performance.
\item Choose the smallest submodel (size) that is sufficiently close to the reference model prediction utility score.
\end{enumerate}
